---
title: "preprocessing_SSP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First of all, we must obtain the year to year values for GDP and Population from the SSP predictions. 

# SSP year to year smoothing

From the SSP database we can obtain 5 year intervals of GDP & population levels. In our case we must obtain the year to year levels.

With this purpose we will assume the data is smooth (as seen in the original SSP paper) and extract the year to year data through a b-spline basis for each country.

First, some preprocessing must be done from the original dataset:
```{r}
library(openxlsx)
library(fda)

ssp  <- read.xlsx("data/ssp_iiasa.xlsx",sheet=1) # Gross available energy, first sheet should be the best, even if in the presentation we chose the fourth
 
north_europe <- c("Denmark","Estonia","Latvia","Lithuania","Finland","Sweden","Norway","Iceland","United Kingdom","Ireland")

east_europe <- c("Bulgaria","Czechia","Hungary","Poland","Romania","Slovakia","Turkey","Ukraine")

south_europe <- c("Greece","Spain","Italy","Cyprus","Malta","Portugal","Croatia","Slovenia","North Macedonia","Albania","Serbia")

central_europe <- c("France","Belgium","Germany (until 1990 former territory of the FRG)","Luxembourg","Netherlands","Austria")

# other EU countries
other_europe     <- c("Montenegro", "Albania", "Bosnia and Herzegovina","Moldova")


ssp_clean_cols <- ssp[,c("SCENARIO","REGION","VARIABLE","UNIT","2010","2015",
                         "2020","2025","2030","2035","2040","2045","2050","2055",
                         "2060","2065","2070","2075","2080","2085","2090","2095",
                         "2100" )]   

country_codes <- data.frame(c(north_europe,east_europe,south_europe,central_europe, other_europe),
                            c("DNK","EST","LVA","LTU","FIN","SWE","NOR","ISL","GBR","IRL", 
                              "BGR","CZE","HUN","POL","ROU","SVK","TUR","UKR",
                              "GRC","ESP","ITA","CYP","MLT","PRT","HRV","SVN","MKD","ALB","SRB",
                              "FRA","BEL","DEU","LUX","NLD","AUT","MNE","ALB","BIH","MDA"))

colnames(country_codes) <- c("Name","Code")

country_name2code <- function(code="",name=""){
    # Returns name and code for a given name or code
    if(nchar(name)>1)
        return(country_codes[country_codes["Name"] == name][2]) #return code
    else if(nchar(code)>1)
        return(country_codes[country_codes["Code"] == code][1]) #return name
} 

ssp_clean_cols[,-1:-4] <- lapply(ssp_clean_cols[,-1:-4], as.numeric) # transform data into numeric
ssp_clean_cols[,"REGION"] <- sapply(ssp_clean_cols[,"REGION"], country_name2code) # Use country names

ssp_eu <- ssp_clean_cols[!is.na(ssp_clean_cols$REGION),] # Discard countries with Na values

# ssp_eu <- rbind(colnames(ssp_eu), ssp_eu) # if wanting to remove colnames and use as rows
# colnames(ssp_eu) <- NULL 

# length(unique(ssp_eu$REGION)) == length(countries.list) # check if data for all countries is present

```

Let us separate the data for GDP and Population.
```{r}
# Divide GDP and Population datasets
ssp_gdp <- ssp_eu[ssp_eu$VARIABLE=="GDP|PPP",] 
ssp_pop <- ssp_eu[ssp_eu$VARIABLE=="Population",]

```

Next, a double loop must be done for each country (and for each dataset) and within  each ocuntry, for each SSP scenario.

```{r}
ssp.year.grid <- seq(2010,2100,5)

ssp.year.grid.str <-c("2010","2015","2020","2025","2030","2035","2040","2045","2050","2055",
                      "2060","2065","2070","2075","2080","2085","2090","2095","2100")

# Create empty df to populate
new.ssp_gdp <- data.frame(matrix(ncol = 4+length(2010:2100), nrow = 0))
new.ssp_pop <- data.frame(matrix(ncol = 4+length(2010:2100), nrow = 0))

for(ssp.var.idx in 1:2){
    if(ssp.var.idx == 1)
        ssp_ <- ssp_gdp
    else
        ssp_ <- ssp_pop
    
    for(sspX in unique(ssp_$SCENARIO)){  # For each scenario
        ssp_sspX <- ssp_[ssp_$SCENARIO==sspX,]
        for(c in 1:length(ssp_sspX$REGION)){ # For each country
            
            ssp_sspX.data <- as.numeric(ssp_sspX[c,ssp.year.grid.str])
            
            # With roahd
            # f_data <- fData(ssp.year.grid, ssp_sspX[,ssp.year.grid.str])
            # plot(f_data)
            
            times_basis <-ssp.year.grid 
            knots    <- c(ssp.year.grid) #Location of knots
            n_knots   <- length(knots) #Number of knots
            n_order   <- 2 # order of basis functions: cubic bspline: order = 3 + 1
            n_basis   <- length(knots) + n_order - 2;
            
            basis <- create.bspline.basis(c(min(times_basis),max(times_basis)),n_basis,n_order,knots)
            # plot(basis)
            
            ys <- smooth.basis(argvals=times_basis, y=ssp_sspX.data, fdParobj = basis)
            
            xfd <- ys$fd
            # plotfit.fd(ssp_sspX.data, times_basis, xfd) #  Plot the curve along with the data
            
            ssp_sspX.yby <- eval.fd(seq(2010,2100,1), xfd) # Obtain data year by year
            
            # plot(seq(2010,2100,1), ssp_ssp1.yby)
            if(ssp.var.idx == 1)
                new.ssp_gdp <- rbind(new.ssp_gdp, c(sspX, ssp_sspX$REGION[c],
                                                 unique(ssp_$VARIABLE)[1],
                                                 unique(ssp_$UNIT)[1],
                                                 ssp_sspX.yby))
            else
                new.ssp_pop <- rbind(new.ssp_pop, c(sspX, ssp_sspX$REGION[c],
                                                 unique(ssp_$VARIABLE)[1],
                                                 unique(ssp_$UNIT)[1],
                                                 ssp_sspX.yby))
        }
    }
}
# rename columns
colnames(new.ssp_gdp) <- c(colnames(ssp_gdp[,1:4]), 2010:2100)    
colnames(new.ssp_pop) <- c(colnames(ssp_pop[,1:4]), 2010:2100)
```



Now we must complete with real values from years previous to the predictions done in the SSP (1990-2010):  
```{r}
# Auxiliary function to convert 2005 dollars to current dollars
to_current_dollars <- function(dollars_2005){
    return(1.37*as.numeric(dollars_2005)) # us bureau of labour statistics (Jan 2005-Jan 2021)
    # return(1.46*dollars_2005) # us bureau of labour statistics (Jan 2005-Dec 2021)
}

# World bank data 
prev.gdp <- read.xlsx("data/UN_GDP_1990-2021.xlsx") # https://data.worldbank.org/indicator/NY.GDP.MKTP.PP.CD?end=2020&locations=EU&start=1990
prev.pop <- read.xlsx("data/UN_Population_1990-2021.xlsx") # https://data.worldbank.org/indicator/SP.POP.TOTL?locations=EU

# Get only 27 eu countries
prev.gdp <- prev.gdp[prev.gdp$Country.Code %in% country_codes$Code,]
prev.pop <- prev.pop[prev.pop$Country.Code %in% country_codes$Code,]

# GDP to billions + population to million
prev.gdp[,c(-1,-2,-3)] <- prev.gdp[,c(-1,-2,-3)]/1000000000
prev.pop[,c(-1,-2,-3)] <- prev.pop[,c(-1,-2,-3)]/1000000

# Use country names
prev.gdp[,"Country.Name"] <- sapply(prev.gdp[,"Country.Code"], country_name2code) 
prev.pop[,"Country.Name"] <- sapply(prev.pop[,"Country.Code"], country_name2code) 
# Only use countries present in current our Data 
prev.gdp <- prev.gdp[prev.gdp$Country.Name %in% unique(new.ssp_gdp$REGION),]
prev.pop <- prev.pop[prev.pop$Country.Name %in% unique(new.ssp_pop$REGION),]
```

REVIEW: Both datasets should be joined together, assuming SSP2 as the most possible scenario ("[...] moderate challenges of both kinds and is intended to represent a future in which development trends are not extreme in either of the dimensions, [...]"), but other more extreme scenarios could be explored: "The central case is not meant to be more likely than any of the other storylines or pathways. In fact, historical development of GHG emissions has often followed trajectories close to the upper bound of the range of earlier emissions scenarios, such as those from SRES":
```{r}
# Select scenario
new.ssp_gdp.ssp.2 <- new.ssp_gdp[new.ssp_gdp == "SSP2",]

new.ssp_pop.ssp.2 <- new.ssp_pop[new.ssp_pop == "SSP2",]

# First we need to transform the SSP GDP data to current dollars + select regions in our data
new.ssp_gdp.ssp.2[,c(-1,-2,-3, -4)] <- lapply(new.ssp_gdp.ssp.2[,c(-1,-2,-3, -4)], to_current_dollars)
new.ssp_gdp.ssp.2 <- new.ssp_gdp.ssp.2[order(new.ssp_gdp.ssp.2$REGION),] # Sort by Country code

new.ssp_pop.ssp.2 <- new.ssp_pop.ssp.2[order(new.ssp_pop.ssp.2$REGION),] # Sort by Country code

```


Combine both datasets
```{r}
ord1=order(prev.gdp$Country.Name)
ord2=order(new.ssp_gdp.ssp.2$REGION)


comb.gdp <- cbind(UNIT=rep("Current International $",
                  dim(new.ssp_gdp.ssp.2)[1]),
                  prev.gdp[ord1,1:23], #valores conocidos
                  new.ssp_gdp.ssp.2[ord2,c(-1:-4)]) #predicciones

ord1=order(prev.pop$Country.Name)
ord2=order(new.ssp_pop.ssp.2$REGION)
comb.pop <- cbind(
                  UNIT=rep("Current International $", dim(new.ssp_pop.ssp.2)[1]),
                  prev.pop[ord1,1:23],
                  new.ssp_pop.ssp.2[ord2,c(-1:-4)])
```

```{r}
write.xlsx(comb.gdp, "data/ssp_iiasa_YtY_gdp.xlsx", overwrite = T)
write.xlsx(comb.pop, "data/ssp_iiasa_YtY_pop.xlsx", overwrite = T)
```


Repeat smoothing procedure to obtain 3 results per year 

```{r}
comb.year.grid <- seq(1990,2100,1)

# Create empty df to populate
new.comb.gdp <- data.frame(matrix(ncol = 4+length(1990:2100), nrow = 0))
new.comb.pop <- data.frame(matrix(ncol = 4+length(1990:2100), nrow = 0))

for(ssp.var.idx in 1:2){
    if(ssp.var.idx == 1)
        ssp_ <- comb.gdp
    else
        ssp_ <- comb.pop 
    
    for(c in 1:length(ssp_$REGION)){ # For each country
        
        ssp_.data <- as.numeric(ssp_[c,as.character(1990:2100)]) 
        
        # With roahd
        # f_data <- fData(ssp.year.grid, ssp_sspX[,comb.year.grid.str])
        # plot(f_data)
        
        times_basis <- comb.year.grid 
        knots       <- c(comb.year.grid) #Location of knots
        n_knots     <- length(knots) #Number of knots
        n_order     <- 2 # order of basis functions: cubic bspline: order = 3 + 1
        n_basis     <- length(knots) + n_order - 2;
        
        basis <- create.bspline.basis(c(min(times_basis),max(times_basis)),n_basis,n_order,knots)
        # plot(basis)
        
        ys <- smooth.basis(argvals=times_basis, y=ssp_.data, fdParobj = basis)
        
        xfd <- ys$fd
        # plotfit.fd(ssp_sspX.data, times_basis, xfd) #  Plot the curve along with the data
        
        ssp_.yby <- eval.fd(seq(1990,2100,1/3), xfd) # Obtain data year by year
        
        # plot(seq(2010,2100,1/3), ssp_ssp1.yby)
        if(ssp.var.idx == 1)
            new.comb.gdp <- rbind(new.comb.gdp, c("SSP2", ssp_$REGION[c],
                                             unique(ssp_$VARIABLE)[1],
                                             unique(ssp_$UNIT)[1],
                                             ssp_.yby))
        else
            new.comb.pop <- rbind(new.comb.pop, c("SSP2", ssp_$REGION[c],
                                             unique(ssp_$VARIABLE)[1],
                                             unique(ssp_$UNIT)[1],
                                             ssp_.yby))
    }
}
# rename columns
colnames(new.comb.gdp) <- c(colnames(comb.gdp[,1:4]), seq(1990,2100,1/3))    
colnames(new.comb.pop) <- c(colnames(comb.pop[,1:4]), seq(1990,2100,1/3))

rownames(new.comb.gdp) <- NULL
rownames(new.comb.pop) <- NULL

```



Let us write the obtained year to year values to a new spreadsheet to use for our predictions:
```{r}
# Write to xlsx
write.xlsx(new.comb.gdp, "data/ssp_iiasa_.3Yt.3Y_gdp.xlsx", overwrite = T)
write.xlsx(new.comb.pop, "data/ssp_iiasa_.3Yt.3Y_pop.xlsx", overwrite = T)

```

